Printing in a Periphery: a Quantitative Study of Finnish Knowledge Production, 1640-1828

### 2017-02-13 inst/examples/20170201-manuscript.Rmd
manuscript figures

## From particular data to global perspective?

**Idea**: Study early modern knowledge production, all of it (1470-1830)

**How?** Combine library catalogues (Fennica, Kungliga, ESTC, CERL HPB database)

**Important**: cleaning up the data (90% of the effort!) and considering how particular data collections (Finnish National Bibliography, English Short-Title Catalogue etc.) have been formed.

**How to fail**: take worldcat or similar combination of datasets and analyse it as a whole without cleaning it up first.

**Overcoming a national gaze** by focus on knowledge production through more complex cross-national network of cities

@@

### Data: bibliographic catalogues 1640-1828

**Fennica** Finnish National bibliography

- `r nrow(subset(df0, catalog == "Fennica" & publication_year >= 1640 & publication_year < 1828))` documents 1640-1828

**Kungliga**  Swedish National bibliography

- `r nrow(subset(df0, catalog == "Kungliga" & publication_year >= 1640 & publication_year < 1828))` documents 1640-1828

**Why this project?**

- Visible roots of the diversity of "Finnishness" reflected in knowledge production, 1640-1828

@@

## Methods: Open data analytical ecosystem

**Customized** our dedicated R packages for bibliographic analysis combine flexibility with automation

**Automated** scales well with new data and unforeseen changes; a million documents cleaned up in 30 minutes

**Open source** fully transparent and reproducible data analytics: https://github.com/rOpenGov/fennica

**Research potential** greatly underestimated; world is full of metadata collections

![Data Science Workflow](data-science.png)  
\tiny{(Figure by Hadley Wickham)}

@@

## Preprocess & Enrich

```{r, message=FALSE, warning=FALSE, echo=FALSE}
library(bibliographica, quiet = TRUE)
```

**Clean up messy entries**

```{r, message=FALSE, warning=FALSE,echo=TRUE}
polish_physical_extent("iii-xxiv, 118, [2] p.")$pagecount
```

**Enrich data** (geocoordinates, gender, ..)

```{r, message=FALSE, warning=FALSE}
get_country("Porvoo")
```

- Parse, clean up, enrich, summarise, analyze, visualize, report..

@@

\begin{center}
\Large{1. General trends and political events}
\end{center}

\begin{center}
\tiny{\textcolor{gray}{General trends \& political events}}
\end{center}

### Turku as a Swedish University Town among Uppsala, Tartu and Stockholm (+ Lund & Greifswald)

```{r map1, echo=FALSE, message=FALSE, warning=FALSE, fig.width=20, fig.height=4, eval=TRUE}
library(dplyr)
library(ggmap)
library(ggplot2)
library(gisfin)
source("funcs.R")
theme_set(get_theme_map())
i <- 1650
mapdata1 <- filter(df0,
              publication_decade == i &
              (catalog == "Kungliga" & 
	            publication_place %in% c("Stockholm", "Uppsala",
	                         "Greifswald", "Lund", "Tartu"))|
	  (catalog == "Fennica" &
	      publication_place == "Turku")) %>%
      group_by(publication_decade, publication_place,
	       latitude, longitude) %>%
      summarise(n = n())

mapdata <- bind_rows(mapdata1, 
  as.data.frame(list(publication_decade = 1650, publication_place = "Lund", 
                     latitude = 55.70293, longitude = 13.19294, n = 0)))

mapdata$size <- mapdata$n
mapdata$highlight <- rep("black", nrow(mapdata))
mapdata$highlight[mapdata$publication_place %in% c("Turku")] <- "red"
mapdata$highlight <- factor(mapdata$highlight)

mapdata$highlight2 <- rep(1, nrow(mapdata))
mapdata$highlight2[mapdata$publication_place %in% c("Greifswald", "Lund")] <- 9
mapdata$highlight2 <- factor(mapdata$highlight2)

region <- c(8, 40, 31, 76)
names(region) <- c("left", "bottom", "right", "top")
#mymap <- get_map(location=geobox(region), color = "bw", source="osm",  maptype="terrain")
#mymap <- get_map(location=geobox(region), source="stamen", maptype = "watercolor")
mymap <- get_map(location=geobox(region), maptype="roadmap")

# Pick the investigated period (sliding window)
# Sum up the years within this sliding window for each element    	    
dfw <- mapdata %>%
    	     group_by(publication_place, latitude, longitude, highlight, highlight2) %>%
		       summarize(n = sum(size))
dfw$size <- log10(1 + dfw$n)
dfw$size[!dfw$highlight2 == 1] <- 0.5*max(dfw$size)

# Plot map
p <- ggmap(mymap) +

        	geom_point(data = subset(dfw, highlight2 == 1),
      	    aes(
      	      x = longitude, 
      	      y = latitude,
	            size = size,
	            shape = 1, 
	            #shape = highlight2, 
	            fill = highlight), 
      	      pch = 21, stroke = 1) +

        	geom_point(data = subset(dfw, !highlight2 == 1),
      	    aes(
      	      thickness = 10,
      	      x = longitude, 
      	      y = latitude,
	            shape = 19, 
	            #shape = highlight2, 
	            color = highlight), 
      	      colour = "black", 
      	      shape = 21, # Point shape 
      	      size = 7,   # Point size
      	      stroke = 1.5) + # Point edge thickness

	        ggtitle(i) +
	        scale_size(range = c(2, 15)) +
	        theme(title = element_text(size = 30), legend.position="none") +
	        scale_fill_manual(values = c("darkgray", "red")) #+
	        #scale_shape_manual(values = c(19,1))
print(p)
```

Map by Stamen

\begin{center}
\tiny{\textcolor{gray}{General trends \& political events}}
\end{center}

### Publishing activity over time

- 1700-1721 Great Northern War & Fennica drop during "the Great Wrath"
- 1760-1770 Pamphleteering plays a crucial role during abolished censorship
- 1803-1815 Napoleonic wars

```{r publishingovertime, echo=FALSE, message=FALSE, warning=FALSE, fig.width=12, fig.height=4, out.width="300px"}
library(magrittr)
library(ggplot2)
library(dplyr)
df2 <- df0 %>% group_by(publication_decade, catalog) %>%
               summarise(n = n())
df2$highlight <- rep(FALSE, nrow(df2))
df2$highlight[df2$catalog == "Kungliga" &
	      df2$publication_decade %in% c(1770, 1810)] <- TRUE
df2$highlight[df2$catalog == "Fennica" &
	      df2$publication_decade %in% c(1710, 1810)] <- TRUE
v <- seq(1500, 1800, 100)
theme_set(theme_bw(20))
p1 <- ggplot(df2, aes(x = publication_decade, y = n)) +
     geom_line(aes(color = catalog), size = 1) +
     geom_point(aes(shape = catalog, color = catalog, size = highlight)) +
     ylab("Title count (n)") + xlab("Publication year") +
     #theme(legend.position = "none") +
     scale_x_continuous(breaks = v, labels = v) +
     scale_color_manual(values = c("blue", "darkgreen")) + 
  	 scale_size_discrete(range = c(2, 8)) +
     guides(color = guide_legend(reverse = TRUE), size = "none", shape = "none")

df <- df0 %>% filter(catalog == "Kungliga")
# Add indicator of the publication period
df$period <- cut(df$publication_year,
           breaks = c(1757, 1766, 1775, 1783),
		 include.lowest = TRUE,
		 right = FALSE
		 )

# Remove entries outside the investigated publication periods
df <- df %>% filter(!is.na(period)) %>%
	     group_by(period) %>%
	     summarise(n = n(),
		  paper = sum(paper, na.rm = TRUE)) %>%
        mutate(paper.per.title = paper/n)

p2 <- ggplot(df,
       aes(x = period, y = paper.per.title)) + 
       geom_bar(stat = "identity") +
       scale_x_discrete(labels = c("1757-1765", "1766-1774", "1775-1783")) +
       ylab("Paper per title") + xlab("")

library(gridExtra)
grid.arrange(p1, p2, nrow = 1)
```

\begin{center}
\tiny{\textcolor{gray}{General trends \& political events}}
\end{center}

### Swedish knowledge production & cycles of representation

An example how historical events, namely **parliamentary assemblies** in the Swedish case, have an effect on knowledge production.

```{r riksdar5, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=5}

minyear <- 1700
  maxyear <- 1800
  catalog <- "Kungliga"

#for (catal in as.character(unique(df0$catalog))) {

for (catal in "Kungliga") {

  df <- subset(dfsel, publication_year >= 1700 & publication_year < 1800 )

  rect_left <- c(min(na.omit(dfm$date)),
               1719-.5, 1719+.5, # Stockholm 20 januari 1719 1 juni 1719
               1734-.5, 1734+.5, # Stockholm 14 maj 1734 14 december 1734
               1765-.5, 1766+.5, # Stockholm 21 februari 1765 21 oktober 1766
               1769-.5, 1770+.5, # Norrköping & Stockholm 22 april 1769 5 februari 1770
               1771-.5, 1772+.5, # Stockholm 19 juni 1771 12 september 1772
               1792-.5, 1792+.5, # Gävle 26 januari 1792 24 februari 1792         
               max(na.omit(dfm$date)))
  rectangles <- data.frame(
    xmin = rect_left[-length(rect_left)],
    xmax = rect_left[-1],
    ymin = ymin,
    ymax = ymax
    )
  rectangles$shade <- rep(c("Background", "Highlight"), length = nrow(rectangles))

 riksplace = c("Stockholm", "Stockholm", "Stockholm", "Norrköping", "Stockholm", "Gävle")
 cols = c("white", "Purple", "yellow", "orange")
 rectangles$shade[rectangles$shade == "Highlight"] = riksplace
 rectangles$shade = factor(rectangles$shade)

  # Draw Figure
  theme_set(theme_bw(20))
  p <- ggplot()
  p <- p + geom_rect(data = rectangles, 
       aes(xmin=xmin, xmax=xmax, ymin=ymin, ymax=ymax, fill=shade), alpha=0.8) + 
         scale_fill_manual(values = cols) # + guides(fill = "none") 
  p <- p + geom_line(data = dfm, aes(x = date, y = Documents), col = "black")
  p <- p + geom_point(data = dfm, aes(x = date, y = Documents), col = "black")
  p <- p + scale_x_continuous(breaks=seq(min(dfm$date), max(dfm$date), 20))
  p <- p + ggtitle("Publishing activity")
  p <- p + ylab("Documents / Year")
  p <- p + xlab("Year")
  p <- p + ggtitle(paste(catal,"Riksplaces"))
  print(p)
}

\begin{center}
\Large{2. Publishers in Sweden and Turku}
\end{center}

\begin{center}
\tiny{\textcolor{gray}{Publishers in Sweden and Turku}}
\end{center}

### Turku in Sweden: proportions for title count and paper in Swedish University towns

@@

\begin{center}
\tiny{\textcolor{gray}{Publishers in Sweden and Turku}}
\end{center}

### Publishers in Swedish University towns (and Finland)

```{r publishers2-finland, echo=FALSE, message=FALSE, warning=FALSE, cache=FALSE, fig.height=7, fig.width=15}
df <- df0
df <- df %>%
        filter(catalog == "Fennica") %>%
	filter(country == "Finland" | publication_place == "Vyborg") %>%
        filter(publication_year >= 1640 & publication_year <= 1828)

@@

# Order by total publishing activity
npub$publication_place <- factor(npub$publication_place, levels = (npub %>% group_by(publication_place) %>% summarise(total = sum(n)) %>% arrange(desc(total)))$publication_place)
npub <- npub %>% arrange(publication_place)

theme_set(theme_bw(20))
p <- ggplot(npub, aes(x = publication_decade, y = n)) +
       geom_bar(stat = "identity", position = "stack",
       		    aes(fill = publication_place)) + 
       xlab("Publication year") +
       ylab("Unique publishers (n)") +
      scale_fill_brewer(palette="Spectral") +       
       guides(fill = guide_legend(reverse = TRUE, title = "")) #+ 
       #ggtitle(paste("Unique publishers in selected publication places (", catalogue, ")"))

@@

\begin{center}
\tiny{\textcolor{gray}{Publishers in Sweden and Turku}}
\end{center}

### Top publishers in Turku/Fennica

```{r publishers4-fennica, echo=FALSE, message=FALSE, warning=FALSE, cache=FALSE, fig.height=6, fig.width=12}
library(dplyr)
place <- "Turku"
catalogue = "Fennica"
df = df0

@@

\begin{center}
\Large{3. Languages and religious texts in Finland}
\end{center}

\begin{center}
\tiny{\textcolor{gray}{Languages \& religious texts in Finland}}
\end{center}

### Different languages in Turku, Stockholm, Uppsala and Lund

  for (myplace in myplaces) {
    p <- ggplot(subset(dff, publication_place == myplace), aes(x = publication_decade, y = f)) +
      geom_bar(position = "stack", stat = "identity", aes(fill = language)) + 
      xlab("Publication year") +
      ylab("Title count frequency (%)") +
      #ggtitle(paste("Languages (", myplace, "/", catalogue, ")", sep = "")) +
      ggtitle(paste(myplace, "/", catalogue)) +
      scale_fill_brewer(palette="Spectral") + 
      guides(fill = guide_legend(reverse = TRUE, title = ""))
    pics[[paste(catalogue, myplace, sep = "-")]] <- p
  }
}

grid.arrange(pics[[1]],
             pics[[8]], pics[[9]], pics[[10]], 
             nrow = 2)

rm(pics)
```

\begin{center}
\tiny{\textcolor{gray}{Languages \& religious texts in Finland}}
\end{center}

### Topic richness per language (publications in Fennica)

p <- ggplot(df, aes(x = publication_year, y = n, color = language)) +
       geom_line() + geom_point() + 
       geom_smooth(aes(fill = language, color = language)) + 
       ylab("Documents (n)") + xlab("Publication year")

\begin{center}
\tiny{\textcolor{gray}{Languages \& religious texts in Finland}}
\end{center}

### Devotional literature (catechisms, hymns, prayers, etc.)

\begin{center}
\tiny{\textcolor{gray}{Languages \& religious texts in Finland}}
\end{center}

### Temporary sermons in Latin and Swedish

- Important role in early modern knowledge production. Language of temporary sermons in contrast to devotional literature has not been analysed in Finland (and Sweden).

\begin{center}
\Large{4. University curriculum and Enlightenment in Finland}
\end{center}

\begin{center}
\tiny{\textcolor{gray}{University curriculum \& Enlightenment in Finland}}
\end{center}

### Dissertation languages in Turku 1640-1828

```{r disslang, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=5}
catalogue <- "Fennica"
langs <- c("Finnish", "Swedish", "Latin", "German",
           "Russian", "French", "Hebrew", "Other")

df <- df0 %>%
        filter(catalog == catalogue) %>%
        filter(publication_year >= 1640 & publication_year <= 1828) %>%
	filter(dissertation)

  # Classify a document to the specifed language
  # If document is assigned with languages, each case is considered
  # so one doc may have multiple entries corresponding to different languages
  # mean(rowSums(df[, lang]) == 1) # 93% Fennica docs have just 1 language
  # Combine data frames for specified languages
  dflsub <- filter(df, df[[lan]])
  if (nrow(dflsub)>0) {
    dflsub$language <- gsub("language.", "", lan)
    dfl <- bind_rows(dfl, dflsub)
  }
}

df <- dfl %>% group_by(publication_decade, language) %>%
     	     summarise(n = n())

# Calculate percentages
dff <- spread(df, language, n, fill = 0)
dff[, -1] <- 100 * t(apply(dff[, -1], 1, function (x) {x/sum(x)}))
dff <- melt(dff, "publication_decade");
colnames(dff) <- c("publication_decade", "language", "f")

# Order by language counts
dff$language <- factor(dff$language,
   levels = (df %>% group_by(language) %>%
                    summarise(total = sum(n)) %>%
		    arrange(total))$language);
dff <- dff %>% arrange(desc(language))

theme_set(theme_bw(20))
p <- ggplot(dff, aes(x = publication_decade, y = f)) +
     geom_bar(position = "stack", stat = "identity", aes(fill = language)) + 
     xlab("Publication year") +
     ylab("Title count frequency (%)") +
     ggtitle(paste("Dissertation languages (", catalogue, ")", sep = "")) +
     scale_fill_brewer(palette="Spectral")
print(p)
```

\begin{center}
\tiny{\textcolor{gray}{University curriculum \& Enlightenment in Finland}}
\end{center}

### Printing by topics in different languages in Turku

```{r topics-103B, echo=FALSE, message=FALSE, warning=FALSE, fig.width=12, fig.height=8}
sel = c("filosofia","luonnonfilosofia","metafysiikka","logiikka")
df = dfl
df$hit <- apply((sapply(sel, function (x) {grepl(x, tolower(df$subject_topic))})), 1, any)  
df <- df %>% filter(hit)
df <- df %>% group_by(publication_decade, language) %>%
             summarise(n = n(),
                 paper = sum(paper, na.rm = TRUE))

# PAPER CONSUMPTION
p1 <- ggplot(df, aes(x = publication_decade, y = paper, group = language)) +
       geom_bar(aes(fill = language), position = "stack", stat = "identity") +                     
       xlab("Publication year") +
       ylab("Paper") +
       guides(fill = guide_legend(reverse = TRUE, title = "")) +
       ggtitle("Philosophy")

sel = c("oppihistoria","antiikki","historia")
# Selected catalogue with selected topics
df = dfl
df$hit <- apply((sapply(sel, function (x) {grepl(x, tolower(df$subject_topic))})), 1, any)  
df <- df %>% filter(hit)
df <- df %>% group_by(publication_decade, language) %>%
             summarise(n = n(),
	               paper = sum(paper, na.rm = TRUE))

# PAPER CONSUMPTION
p2 <- ggplot(df, aes(x = publication_decade, y = paper, group = language)) +
       geom_bar(aes(fill = language), position = "stack", stat = "identity") +             
       xlab("Publication year") +
       ylab("Paper") +
       guides(fill = guide_legend(reverse = TRUE, title = "")) +
       ggtitle("History")

sel =   c("psykologia","yhteiskuntafilosofia","valtiofilosofia","talous","raha","moraali","velvollisuudet","kasvatus","maanviljely","maatalous","kalastus")

# Selected catalogue with selected topics
df = dfl
df$hit <- apply((sapply(sel, function (x) {grepl(x, tolower(df$subject_topic))})), 1, any)  
df <- df %>% filter(hit)
df <- df %>% group_by(publication_decade, language) %>%
             summarise(n = n(),
	               paper = sum(paper, na.rm = TRUE))

# PAPER CONSUMPTION
theme_set(theme_bw(20))
p3 <- ggplot(df, aes(x = publication_decade, y = paper, group = language)) +
       geom_bar(aes(fill = language), position = "stack", stat = "identity") +              
       xlab("Publication year") +
       ylab("Paper") +
       guides(fill = guide_legend(reverse = TRUE, title = "")) +
       ggtitle("Economy")
grid.arrange(p1, p2, p3, nrow = 3)
```

\begin{center}
\tiny{\textcolor{gray}{University curriculum \& Enlightenment in Finland}}
\end{center}

### Rise of the octavo sized book as vehicle of Enlightenment

```{r LIBER-13, echo=FALSE, message=FALSE, warning=FALSE, echo=FALSE, fig.width=11, fig.height=9, fig.show="hold", out.width="160px"}
for (catal in unique(df0$catalog)) {
  df2 <- subset(df0, catalog == catal) %>% group_by(publication_decade, gatherings) %>% summarize(paper = sum(paper, na.rm = TRUE), n = n()) 
  df2 <- filter(df2, gatherings %in% setdiff(names(which(table(df2$gatherings) >= 15)), "NA"))
  df2$highlight <- rep("Other", nrow(df2))
  df2$highlight[df2$gatherings == "8vo"] <- "Octavo"

##Thanks !

Institutional support:

- Academy of Finland
- University of Helsinki
- Digitalia / The Regional Council of South Savo

Data providers:

- National Library of Finland
- National Library of Sweden

Code infrastructure:

- Github
- R project

For more info, see:

- \textcolor{blue}{https://github.com/rOpenGov/fennica}

\begin{figure}
\begin{tabular}{cccccc}
\includegraphics[width=2cm,height=2cm]{aka.png}&
\includegraphics[width=1cm,height=1cm]{HY-CC.png}&
\includegraphics[width=1cm,height=1cm]{kansalliskirjasto.png}&
\includegraphics[width=1cm,height=1cm]{kungliga.png}&
\includegraphics[width=1cm,height=1cm]{github-logo.png}&
\includegraphics[width=1cm,height=1cm]{Rlogo.png}
\end{tabular}
\end{figure}

### Duplicate publications in Fennica/Kungliga

```{r duplicates, echo=FALSE, message=FALSE, warning=FALSE, cache=FALSE, fig.height=5, fig.width=10}
df <- df.combined.preprocessed %>%
        filter(publication_year >= 1640 & publication_year <=1828) %>%
        group_by(catalog, publication_decade) %>%
	summarise(duplicates = 100 * mean(remove))

p <- ggplot(df, aes(x = publication_decade, y = duplicates, fill = catalog)) +
       geom_bar(position = "dodge", stat = "identity") +
       ylab("Duplicates (%)") +
       scale_fill_manual(values = c("blue", "red")) 
       #ggtitle(paste("Certain duplicates Fennica-Kungliga", paste(range(na.omit(df$publication_decade)), collapse = "-")))

>>>>

### Fig. 1: Number of publication places per decade

```{r places_per_decade, message=FALSE, warning=FALSE, fig.width=10, fig.height=3, echo=FALSE}
library(dplyr)
library(ggplot2)
theme_set(theme_bw(20))
df <- df0 %>% filter(publication_year >=1640 & publication_year <=1828) %>%
              group_by(catalog, publication_decade) %>%
	      summarise(n = length(unique(na.omit(publication_place))))
p <- ggplot(df, aes(x = publication_decade, y = n, color = catalog)) + 
       geom_point() + 
       geom_line() + 
       ylab("Publication places (n)") + xlab("Publication year") +
       scale_color_manual(values = c("blue", "darkgreen")) + 
       guides(color = guide_legend(reverse = TRUE, title = "")) +
       ggtitle("Number of unique publication places")
print(p)
```

@@

### Fig 2: Number of documents per decade

@@

### Fig 3: Rise of the octavo sized book as vehicle of Enlightenment

```{r LIBER-13, echo=FALSE, message=FALSE, warning=FALSE, echo=FALSE, fig.width=11, fig.height=9, fig.show="hold"}
for (catal in unique(df0$catalog)) {
  df2 <- subset(df0, catalog == catal) %>% group_by(publication_decade, gatherings) %>% summarize(paper = sum(paper, na.rm = TRUE), n = n()) 
  df2 <- filter(df2, gatherings %in% setdiff(names(which(table(df2$gatherings) >= 15)), "NA"))
  df2$highlight <- rep("Other", nrow(df2))
  df2$highlight[df2$gatherings == "8vo"] <- "Octavo"

@@

### Fig 4: Average title length (word count)

```{r title_length_by_lang, echo=FALSE, message=FALSE, warning=FALSE, cache=FALSE, fig.height=7, fig.width=15}
dfo <- df0

# Calculate title length (word count)
dfo$title_wordcount <- sapply(strsplit(as.character(dfo$title), " "), length)
dfo$title_wordcount[is.na(dfo$title)] <- NA

top.language <- setdiff(names(top(dfo, "language", 11)), "NA")

@@

### Fig 5: Turku in Sweden: proportions for title count and paper in Swedish University towns

@@

### Fig 6: Publishing activity over time

```{r publishingovertime, echo=FALSE, message=FALSE, warning=FALSE, fig.width=12, fig.height=4, out.width="300px"}
library(magrittr)
library(ggplot2)
library(dplyr)
v <- seq(1500, 1800, 100)
theme_set(theme_bw(20))
df <- df0 %>% filter(catalog == "Kungliga")

# Add indicator of the publication period
df$period <- cut(df$publication_year,
           breaks = c(1757, 1766, 1775, 1783),
		 include.lowest = TRUE,
		 right = FALSE
		 )

# Remove entries outside the investigated publication periods
df.all <- df %>% filter(!is.na(period)) %>%
	     group_by(period) %>%
	     summarise(n = n(),
		  paper = sum(paper, na.rm = TRUE)) %>%
        mutate(paper.per.title = paper/n)

# Remove entries outside the investigated publication periods
df.cities <- df %>% filter(!is.na(period)) %>%
	     group_by(period, publication_place) %>%
	     summarise(n = n(),
		  paper = sum(paper, na.rm = TRUE)) %>%
        mutate(paper.per.title = paper/n)

df.all$publication_place <- "All"
df <- bind_rows(df.all, df.cities) %>%
        filter(publication_place %in% c("Stockholm", "Uppsala", "Lund", "All"))

p <- ggplot(df,
       aes(fill = period, y = paper.per.title, x = publication_place)) + 
       geom_bar(stat = "identity", color = "black", position = "dodge") +
       #scale_x_discrete(labels = c("1757-1765", "1766-1774", "1775-1783")) +
       guides(fill = guide_legend(title = "Period")) +
       ylab("Paper (sheets)") + xlab("") +
       ggtitle("Paper consumption per title")

### Fig. 7: Riksdagar Kungliga

```{r riksdar_kungliga, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=5}

minyear <- 1700
maxyear <- 1800
catal <- "Kungliga"

  df <- subset(dfsel, publication_year >= 1700 & publication_year < 1831 )

  rect_left <- c(min(na.omit(dfm$date)),
               1710-.5, 1710+.5, #
               1713-.5, 1714+.5, #

               1719-.5, 1719+.5, # Stockholm 20 januari 1719 1 juni 1719
               1734-.5, 1734+.5, # Stockholm 14 maj 1734 14 december 1734
               1765-.5, 1766+.5, # Stockholm 21 februari 1765 21 oktober 1766
               1769-.5, 1770+.5, # Norrköping & Stockholm 22 april 1769 5 februari 1770
               1771-.5, 1772+.5, # Stockholm 19 juni 1771 12 september 1772

               1778-.5, 1779+.5, #
               1786-.5, 1786+.5, #
               1789-.5, 1789+.5, #

               1792-.5, 1792+.5, # Gävle 26 januari 1792 24 februari 1792

               1800-.5, 1800+.5, #
               1809-.5, 1810+.5, #
               1812-.5, 1812+.5, #
               1815-.5, 1815+.5, #
               1817-.5, 1818+.5, #
               1823-.5, 1823+.5, #
               1828-.5, 1830+.5, #

               max(na.omit(dfm$date)))
  rectangles <- data.frame(
    xmin = rect_left[-length(rect_left)],
    xmax = rect_left[-1],
    ymin = ymin,
    ymax = ymax
    )

  rectangles$shade <- rep(c("Background", "Highlight"), length = nrow(rectangles))

  riksplace <- c(rep("Unknown", 2), "Stockholm", "Stockholm", "Stockholm", "Norrköping", "Stockholm", rep("Unknown", 3), "Gävle", rep("Unknown", 7))
  #cols <- c("white", "Purple", "yellow", "orange", "darkgray")
  cols <- c("white", "gray", "darkgray")
  rectangles$place <- rectangles$shade
  rectangles$place[rectangles$place == "Highlight"] <- riksplace
  rectangles$place[rectangles$place == "Background"] <- "" 
  rectangles$shade <- factor(rectangles$shade)

  # Draw Figure
  theme_set(theme_bw(20))
  p <- ggplot() +
         geom_rect(data = rectangles, 
           aes(xmin=xmin, xmax=xmax,
	       ymin=ymin, ymax=ymax,
	       fill=shade), alpha=0.8) +
         scale_fill_manual(values = cols) +
	 guides(fill = "none")

  p <- p + geom_text(data = rectangles, angle = 90, aes(x=xmin + .5, y=.9*ymax, label=place), size = 6)

  p <- p + geom_line(data = dfm, aes(x = date, y = Documents), col = "black")
  p <- p + geom_point(data = dfm, aes(x = date, y = Documents), col = "black")
  p <- p + scale_x_continuous(breaks=seq(min(dfm$date), max(dfm$date), 10))
  p <- p + ggtitle("Publishing activity")
  p <- p + ylab("Documents / Year")
  p <- p + xlab("Year")
  p <- p + ggtitle(paste(catal,"Riksplaces"))
  print(p)

### Fig. 8: Riksdagar Fennica

```{r riksdar_fennica, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=5}
minyear <- 1760
maxyear <- 1860
catal <- "Fennica"

df00 <- df.full %>% filter(publication_year >= minyear & publication_year <= maxyear)

  # Use timeinterval year intervals
  dfsel <- subset(df00, catalog == catal)

  df <- df00
  df <- subset(df, catalog == catal)
  df <- subset(df, publication_year >= minyear & publication_year <= maxyear)
  df <- subset(dfsel, publication_year >= minyear & publication_year <= maxyear)

  df$unity <- rep(1, nrow(df))
  publications <- tapply(df$unity, list(df$timeunit), sum)

  publications[is.na(publications)] <- 0 # Set NAs to 0
  publications <- publications/timeinterval # Instead of decadal sum, use average annual output 
  dfm <- melt(publications) 
  names(dfm) <- c("Time", "Documents")
  dfm <- transform(dfm, date = as.numeric(as.character(Time)))
  ymin <- min(dfm$Documents)
  ymax <- max(dfm$Documents)

  rect_left <- c(min(na.omit(dfm$date)),
               1710-.5, 1710+.5, #
               1713-.5, 1714+.5, #

               1719-.5, 1719+.5, # Stockholm 20 januari 1719 1 juni 1719
               1734-.5, 1734+.5, # Stockholm 14 maj 1734 14 december 1734
               1765-.5, 1766+.5, # Stockholm 21 februari 1765 21 oktober 1766
               1769-.5, 1770+.5, # Norrköping & Stockholm 22 april 1769 5 februari 1770
               1771-.5, 1772+.5, # Stockholm 19 juni 1771 12 september 1772

               1778-.5, 1779+.5, #
               1786-.5, 1786+.5, #
               1789-.5, 1789+.5, #

               1792-.5, 1792+.5, # Gävle 26 januari 1792 24 februari 1792

               1800-.5, 1800+.5, #
               1809-.5, 1810+.5, #
               1812-.5, 1812+.5, #
               1815-.5, 1815+.5, #
               1817-.5, 1818+.5, #
               1823-.5, 1823+.5, #
               1828-.5, 1830+.5, #

               max(na.omit(dfm$date)))

  rectangles <- data.frame(
    xmin = rect_left[-length(rect_left)],
    xmax = rect_left[-1],
    ymin = ymin,
    ymax = ymax
    )

  rectangles$shade <- rep(c("Background", "Highlight"),
  		          length = nrow(rectangles))

  riksplace <- c(rep("Unknown", 2), "Stockholm", "Stockholm", "Stockholm", "Norrköping", "Stockholm", rep("Unknown", 3), "Gävle", rep("Unknown", 7))
  cols <- c("white", "gray", "darkgray")
  rectangles$place <- rectangles$shade
  rectangles$place[rectangles$place == "Highlight"] <- riksplace
  rectangles$place[rectangles$place == "Background"] <- "" 
  rectangles$shade <- factor(rectangles$shade)

  # Draw Figure
  theme_set(theme_bw(20))
  p <- ggplot() +
         geom_rect(data = rectangles, 
           aes(xmin=xmin, xmax=xmax,
	       ymin=ymin, ymax=ymax,
	       fill=shade), alpha=0.8) +
         scale_fill_manual(values = cols) +
	 guides(fill = "none")

  p <- p + geom_text(data = rectangles, angle = 90, aes(x=xmin + .5, y=.9*ymax, label=place), size = 6)

  p <- p + geom_line(data = dfm, aes(x = date, y = Documents), col = "black")
  p <- p + geom_point(data = dfm, aes(x = date, y = Documents), col = "black")
  p <- p + scale_x_continuous(breaks=seq(min(dfm$date), max(dfm$date), 10))
  p <- p + ggtitle("Publishing activity")
  p <- p + ylab("Documents / Year")
  p <- p + xlab("Year")
  p <- p + ggtitle(paste(catal))
  p <- p + xlim(c(minyear, maxyear))
  print(p)
```

### Fig 9A: Fennica - Publishers in Swedish University towns (and Finland)

```{r publishers2-finland, echo=FALSE, message=FALSE, warning=FALSE, cache=FALSE, fig.height=7, fig.width=15}
df <- df0
catal <- "Fennica"
df <- df %>%
        filter(catalog == catal) %>%
	filter(country == "Finland" | publication_place == "Vyborg") %>%
        filter(publication_year >= 1640 & publication_year <= 1828)

@@

# Order by total publishing activity
npub$publication_place <- factor(npub$publication_place, levels = (npub %>% group_by(publication_place) %>% summarise(total = sum(n)) %>% arrange(desc(total)))$publication_place)
npub <- npub %>% arrange(publication_place)
npub$publication_place <- factor(npub$publication_place, levels = rev(unique(npub$publication_place)))

theme_set(theme_bw(20))
p <- ggplot(npub, aes(x = publication_decade, y = n)) +
       geom_bar(stat = "identity", position = "stack",
       		    aes(fill = publication_place)) + 
       xlab("Publication year") +
       ylab("Unique publishers (n)") +
       scale_fill_brewer(palette="Spectral") +       
       guides(fill = guide_legend(reverse = FALSE, title = "")) +
       ggtitle(catal)       
       #ggtitle(paste("Unique publishers in selected publication places (", catalogue, ")"))

print(p)

### Fig 9B: Kungliga - Publishers in Swedish University towns (and Finland)

```{r kungliga9b, echo=FALSE, message=FALSE, warning=FALSE, cache=FALSE, fig.height=7, fig.width=15}
df <- df0
catal <- "Kungliga"
df <- df %>%
        filter(catalog == catal) %>%
	filter(country == "Finland" | publication_place == "Vyborg") %>%
        filter(publication_year >= 1640 & publication_year <= 1828)

# Order by total publishing activity
npub$publication_place <- factor(npub$publication_place, levels = (npub %>% group_by(publication_place) %>% summarise(total = sum(n)) %>% arrange(desc(total)))$publication_place)
npub <- npub %>% arrange(publication_place)
npub$publication_place <- factor(npub$publication_place, levels = rev(unique(npub$publication_place)))

theme_set(theme_bw(20))
p <- ggplot(npub, aes(x = publication_decade, y = n)) +
       geom_bar(stat = "identity", position = "stack",
       		    aes(fill = publication_place)) + 
       xlab("Publication year") +
       ylab("Unique publishers (n)") +
       scale_fill_brewer(palette="Spectral") +       
       guides(fill = guide_legend(reverse = FALSE, title = "")) +
       ggtitle(catal)
       #ggtitle(paste("Unique publishers in selected publication places (", catalogue, ")"))

print(p)

### Fig 10: Publications in Vaasa in Fennica, 1750-1828, with percentage of all books that were not published by the printing house of the court

```{r publishers4-fennica, echo=FALSE, message=FALSE, warning=FALSE, cache=FALSE, fig.height=6, fig.width=12}
library(dplyr)
catalogue = "Fennica"
place <- "Vaasa"
minyear <- 1750
maxyear <- 1828

df <- df0 %>% filter(catalog == catalogue & publication_place == place) %>%	
        filter(publication_year >= minyear & publication_year <= maxyear)

p <- ggplot() + ggtitle("Fig 10: to be added")

# TITLE COUNT
#theme_set(theme_bw(20))
#p <- ggplot(npub, aes(x = publication_decade, y = publisher)) +
#      geom_point(aes(size = n)) + 
#      #geom_line(aes(color = publisher), size = 2) +
#      #geom_smooth(aes(fill = publisher)) +
#      # geom_bar(stat = "identity", position = "stack", aes(fill = publisher)) #+ 
#       xlab("Year") +
#       ylab("") +
#       guides(size = guide_legend(title = "Title count (n)"))
#       #ggtitle(paste("Title count per publisher (", place, "/", catalogue, ")"#, sep = ""))

@@

### Fig 11: Top publishers in Turku/Fennica

```{r publishers-fennica-11, echo=FALSE, message=FALSE, warning=FALSE, cache=FALSE, fig.height=6, fig.width=12}
library(dplyr)
place <- "Turku"
catalogue = "Fennica"
df = df0

@@

### IMAGE 12: Title count per capita
